### 2023-02-22 text/ManuscriptDraft.Rmd
Editions to manuscript

---
new_session: FALSE
#title: |
#  **Detecting small objects and many of them: Using computer vision and deep learning to automate counting of collembola in petri dish samples**
#date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document: default
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
  html_document:
    df_print: paged
fontfamily: mathpazo
fontsize: 12pt
geometry: margin = 1in
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{dcolumn}
- \usepackage{caption}
- \usepackage{float}
- \usepackage{afterpage}
- \usepackage{siunitx}
- \usepackage{amsmath}
#keywords: X
bibliography: ./library.bib
csl: ./journal-of-ecology.csl
editor_options: 
  markdown: 
    wrap: 72
---

```{r eval = FALSE, echo=FALSE}
library(knitr)
knit('ManuscriptDraft.Rmd')
rmarkdown::render("ManuscriptDraft.Rmd", output_file = "ManuscriptDraft.docx")
system2("open","manuscriptDraft.docx")
```

```{r echo=FALSE}
# Function for creating word comments in Rmarkdown
# Use: This text contains a `r word_comment("This is the comment", "comment.")`.

word_comment <- function(comment, highlight = "") {
  if (isTRUE(knitr:::pandoc_to() == "docx")) {
    paste0('[', comment, ']{.comment-start id="0" author="Hjalte M."',
           'date="2020-01"}', highlight, '[]{.comment-end id="0"}')
  }
}
```

Title:

**Detecting small objects and many of them: Computer vision and deep
learning enables high-throughput counting of collembola in petri dish
samples**

#

Authors:

Hjalte M. R. Mann^1^, James J. Scott-Fordsmand^1^, Toke T. Høye^1^

#

Author affiliations:

1: Department of Ecoscience, Aarhus University, C.F. Møllers Allé 4-8,
8000 Aarhus C, Denmark\

#

Corresponding author contact information:

Hjalte Mads Rosenstand Mann\
E-mail: [mann\@ecos.au.dk](mailto:mann@ecos.au.dk){.email}\
Phone: +45 31178585

#

ORCID IDs:

HMRM: 0000-0002-4768-4767, JJSF: 0000-0002-2260-1224, TTH:
0000-0001-5387-3284

#

Keywords: Collembola, computer vision, deep learning

#

Article type: Research article

Word count: 0

\newpage

# Abstract

Requires manual counting of

Recent advances in computer vision and deep learning methods enable
automation of image analysis including detection of objects of interest.

Here's the abstract.

\newpage

# Introduction

Ecotoxiological laboratory studies on soil invertebrates deliver
important assessments of compound toxicities.

Manual live counts problematic because of movement of individuals.
Counting based on images can alleviate this problem. Images also allow
validation of counts.

-Why do we need to count collembola in petri dishes -What is the problem
with the established method -Developments in computer vision -How can
computer vision help solve our problem -What are we presenting

Counting of collembola in petri dish samples in ecotoxicology
experiments requires substantial manual labor by trained technicians.
The process is time-consuming while the result is a simple census of the
present individual collembola. Here, we investigated the feasibility of
automating collembola counting based on a single image of a sample taken
with a handheld camera and a deep learning object detection model.

Object detection models have gained significant attention for their
ability to locate specific objects of interests in images. However,
reliable detection of very small and very large numbers of objects
remains a challenging task. In this regard, the present task is
particularly challenging owing to the very small size of objects, low
distinguishability of animals from background noise, large variation in
number of objects in an image, visual variation between samples, and
variation in the conditions in which the image was taken, how the image
was taken and the camera that was used.

In order to ease the manual counting of animals in an image, it is often
done by taking an image of the sample and performing the counting in
image processing software. This allows for marking of counted individual
to prevent double-counting and missed individuals and validation of
result.

One approach has been semi-automated background removal to create a
binary blob image ideally containing only animals and all the animals in
a dish. This has allowed for subsequent automated counting of blobs and
thus individuals. However, this process still requires substantial
manual effort to clean the image and separate individuals to produce
reliable results.

Ideally, counting would be entirely automated with just a single RGB
image as input.

The basis of this was the output of manual processing of images

We publish the code together with the trained model to facilitate others
to implement automated analysis of ecotoxicology samples.

# Material and methods

-   Overview of method
-   Standard method
-   Translating binary to bounding box format for model training
-   Splitting image/annotation data into training, validation, and
    testing sets
-   Slicing images/annotations for full resolution training
-   Training and testing neural network with sliding window approach

The method involved several steps: Translating binary filter images to
bounding box format; splitting images into training, validation and
testing sets; slicing images and associated annotations to a standard
size in order to not loose information from large images because of
downscaling; training and testing the neural network with a sliding
window approach.

## Generating training and testing data

Images had substantial variation in size, with the smallest image
included being XxX and the largest XxX. We sliced to 640 x 640. Images
were only sliced on dimensions larger than 640, i.e., an image with size
100x1280 would be sliced to two image of 100x640.

Training and testing object detection models requires annotated training
and testing data. Specifically, images containing the objects of
interest together with associated annotation data containing coordinates
for bounding boxes delimiting the objects are required. Here, generation
of annotation data was done through automated processing of existing
outputs from human analysis. The human analysis results in a binary mask
output of each petri dish image with black blobs delimiting animals. To
convert these to bounding boxes, we applied automated contour detection
and bounding box creation using Python OpenCV. The contour detection
produces some errors when there are many objects very close to each
other. Although the error rate seems low, for this test, images with a
very high number of objects were discarded. Images from four tests were
used (Test A-D). Only one image of each sample was used (i.e. for A1.1,
A1.2, etc., only one image was included).

Contour detection and bounding box derivation was performed for a total
of 127 images resulting in a total of 110.553 objects detected and
annotated. See table x for distribution of images and objects between
groups.

Group Images Objects A 44 66056 B 39 17479 C 16 17059 D 28 9959

To decrease risk of objects not contained in bounding box, we expanded
the widths and heights of boxes with 10%.

## Splitting data into training, validation, and testing

The images varied substantially in the number of annotated objects they
contained. For each experimental group, we split the images into
training, validation, and testing at a ratio of 70/15/15 at the level of
images while optimizing for a ratio of 70/15/15 at the level of objects.
If 70% of the images resulted in an unequal remainder or a decimal
value, we rounded to a number that would ensure an equal remainder. This
was done so that we could ensure an equal split between validation and
testing in terms of number of images. We assigned images to groups by
treating the problem as a linear programming problem with the number of
images assigned to each group constrained to the 70/15/15 split while
the distribution of images to groups were optimized towards lowest
deviation from the 70/15/15 split at the level of objects assigned to
each group.

## Slicing images and annotations

As the suggested method will produce overlap between sliced images when
full 640x640 slices cannot be produced, individual animals will
sometimes be present in more than one slice.

Slicing produced the following distribution of sliced images and
objects:

Set Images Objects Training\
Validation\
Testing

TO DO \# images \# objects Train 89\
Val 19\
Test 19\
Total 127

## Training and testing neural network with sliding window approach

Just do the slice+SAHI approach and compare with standard 640x640.

Deep learning methods To test the applicability of object detection
models for automatic detection and counting of collembola, we tested two
model sizes of the object detection architecture YOLOv5: the small
YOLOv5s which takes images with dimensions of 640x640 as input
(hereafter referred to as YOLOv5 640), and the YOLOv5s6 which takes
images of dimensions 1280x1280 (hereafter referred to as YOLOv5 1280).
Additionally, we tested the applicability of the Slicing Aided Hyper
Inference (SAHI) method for each model. This method slices each image
during inference in a sliding window manner and performs inference on
each slice. The method is particularly relevant for detection of small
objects. Training length was set to 1000 epochs with early stopping with
a patience of 100 epochs. Confidence threshold for prediction set to
0.1. Non-maximum suppression IOU threshold set to 0.45.

Evaluating detection performance

Since YOLOv5+SAHI has a different output format than YOLOv5 on its own,
customs scripts were created to 1) translate YOLOv5 output pickle files
to standard YOLOv5 text files and 2) calculate detection performance
variables (precision and recall) for both methods. Precision is the
number of correct predictions out of all predictions made. Recall is the
proportion of the true number of objects that were detected by the
model. F1 is the weighted average of precision and recall. A value of
0.3 was used as the IOU threshold for whether bounding boxes overlapped
sufficiently to be considered correct.

## Resolution per petri dish

To investigate the influence of image resolution on detection
performance, we measured the number of pixels within the limits of the
petri dish in each image.

# Results

-   Overview of original number of images, lowest resolution, highest
    resolution,
-   Translating binary filters to bounding box format o lowest number of
    annotations, highest number of annotations
-   Splitting images+annotations into training, validation, testing,
    with linear optimization (at image level but on condition of number
    of annotations in each set)
-   Slicing images and annotations to avoid downscaling leverage full
    resolution information o Resulting number of images (mean number of
    annotations, sd?)
-   Training and testing model o Prediction accuracy

### Standard approach

The model trained on the standard images reached best validation
accuracy at epoch xxx.

### Slicing

The model trained on the sliced images reached best validation accuracy
at epoch xxx.

Running inference on image slices with

Table 1 Detection accuracy with the different models. Method (IOU 0.3)
Precision Recall F1 YOLOv5 640 0.46 0.19 0.27 YOLOv5 640 + SAHI (128)
0.60 0.12 0.20 YOLOv5 1280 0.54 0.56 0.55 YOLOv5 1280 + SAHI (128) 0.63
0.41 0.50

Fig. 2 Example image from test data set. Green bounding boxes are ground
truth annotations derived from the binary images. Red boxes are
detections made by the YOLOV5 1280 model + SAHI. In this case, most
animals marked in the ground truth are detected by the model (high
recall), but the model produces many false positives (low precision).
Note the overlapping detections in one place (middle of image).

Fig. 3 Example outputs from the two different model sizes with and
without application of SAHI (128x128 window size). P: precision, r:
recall.

Table 3 Variation in performance between images (YOLOv5 1280 + SAHI
128).

Image Precision Recall

# Discussion

-   What did we present
-   Who can use it -- and how o Model ready to test o Code available for
    training on new data and adaption
-   High resolution images taken in a standardized setting for training
    and production likely to improve results and should be considered in
    future studies.
-

The improved precision with increased image dimensions and with the
application of SAHI underlines that a main challenge of the task of
detecting collembola is the very small object sizes relative to the full
image.

Overlapping detections The models tended to produce overlapping
detections. This will have some influence on the accuracy metrics, but
it can be assumed to be minor, since the overlapping detections seems to
be rare. Overlapping detections are usually removed by non-maximum
suppression. However, for the present case with very small object, the
NMS IOU threshold was likely set too high,

Accuracy of the ground truth annotations Considering the similarity
between background and animals, there is likely some level of error in
the annotations. On top of this, the automated extraction of bounding
boxes from the binary images could introduce error, e.g. when there are
many objects close to each other (all though in the present case, such
images were discarded). If the accuracy of these ground truth
annotations is in fact low, it will have detrimental effects on model
accuracy. To investigate this issue it could be relevant to test the
variation in manual annotations between annotators.

Variation in accuracy between test images The model accuracy generally
varied substantially between images within the test set. This is likely
because the model has not been trained on a training set of sufficient
size and variation, which means the models fails to generalize between
images.

Different issues with data means that the present experiment was carried
out on a limited set of images. Issues/challenges with the available
data - Contour errors when objects are many and close. - Test E folder
contains no original images (only b/w), so these cannot be used. - Many
original images without associated b/w image. - Several versions of some
b/w images. Used latest version in each case. - Images have varied
dimensions. Standardized by padding as described. This also means
variation in size of objects (in pixels).

# Acknowledgements

xxx

# Author contributions

HRMR, JJSF and TTH conceived the idea. HMRM conceived and developed the
method with contributions from JJSF and TTH. HMRM wrote all code. HMRM
wrote the manuscript with contributions from JJSF and JJSF. All authors
gave final approval for publication.

# Data availability

The code and data that supports the results in this paper is openly
available at <https://github.com/TECOLOGYxyz/springtail>.

\newpage

# References